

```{r}
library(tidyverse)
library(caret)

dir = "directory/to/data/"

```
By Camilla Uhlbäck


Read in the dataframes
```{r}

# All nouns and ratings from the original dataset
valar <- read.table(paste0(dir,"dfs_vectors/valar.csv"), sep=",", header=T) 

# word embedding files
VH <- read.table(paste0(dir,"dfs_vectors/VH_final.csv"), sep=",", header=T)
VL <- read.table(paste0(dir,"dfs_vectors/VL_final.csv"), sep=",", header=T)
AH <- read.table(paste0(dir,"dfs_vectors/AH_final.csv"), sep=",", header=T)
AL <- read.table(paste0(dir,"dfs_vectors/AL_final.csv"), sep=",", header=T)


# image embedding files
VH.eff <- read.table(paste0(dir,"dfs_vectors/VH_effnet.csv"), sep=",", header=T)
VL.eff <- read.table(paste0(dir,"dfs_vectors/VL_effnet.csv"), sep=",", header=T)
AH.eff <- read.table(paste0(dir,"dfs_vectors/AH_effnet.csv"), sep=",", header=T)
AL.eff <- read.table(paste0(dir,"dfs_vectors/AL_effnet.csv"), sep=",", header=T)

VH.res <- read.table(paste0(dir,"dfs_vectors/VH_resnet.csv"), sep=",", header=T)
VL.res <- read.table(paste0(dir,"dfs_vectors/VL_resnet.csv"), sep=",", header=T)
AH.res <- read.table(paste0(dir,"dfs_vectors/AH_resnet.csv"), sep=",", header=T)
AL.res <- read.table(paste0(dir,"dfs_vectors/AL_resnet.csv"), sep=",", header=T)

```


Testing correlation between valence and arousal, 
starting with normality test to proceed to the correct cor.test
```{r}
# correlation test for valence and arousal

shapiro.test(AH$A.Mean.Sum) # test is significant: the distribution is non-normal
shapiro.test(AL$A.Mean.Sum) # non-normal
shapiro.test(VH$V.Mean.Sum) # non-normal
shapiro.test(VL$V.Mean.Sum) # non-normal

```

Distribution was non-normal, using Spearman's correlation coefficient
```{r}
# spearman's is the nonparametric correlation test

cor.test(VH$V.Mean.Sum, VH$A.Mean.Sum, method = "spearman")   # high valence  
# a low rho (0.26)
cor.test(VL$V.Mean.Sum, VL$A.Mean.Sum, method = "spearman")   # low valence
# low rho (-0.26)
cor.test(AH$V.Mean.Sum, AH$A.Mean.Sum, method = "spearman")   # high arousal
# even worse rho (-0.1)
cor.test(AL$V.Mean.Sum, AL$A.Mean.Sum, method = "spearman")   # low arousal
# the worst rho (-0.08)

# very low or no correlation between the variables valence and arousal
```


**1. Normalisation function**


```{r}

normalise_embedding <- function(df, starts_w){      # function for normalisation
  embedding_lengths <- df %>%
    select(Word, starts_with(!!starts_w)) %>%
    pivot_longer(!Word) %>%
    group_by(Word) %>% 
    mutate(val_2 = value^2) %>%
    summarise(len = abs(sqrt(sum(val_2))))
    
  df %>%
    select(Word, starts_with(!!starts_w)) %>%
    pivot_longer(!Word) %>%
    group_by(Word) %>%
    left_join(embedding_lengths, by = "Word") %>%
    mutate(norm_value = value / len) %>%
    select(Word, name, norm_value) %>%
    pivot_wider(names_from = name, values_from = norm_value) %>% 
    ungroup() %>%
    return()
}

check_normalised <- function(df, starts_w){    # test if length equals 1
  embedding_lengths <- df %>%
    select(Word, starts_with(!!starts_w)) %>%
    pivot_longer(!Word) %>%
    group_by(Word) %>% 
    mutate(val_2 = value^2) %>%
    summarise(len = abs(sqrt(sum(val_2))))
  return(all(near(embedding_lengths$len,1)))
}

```

**2. K-fold results function**

Home made RMSE, R2, MAE, rho; 
showing the correlation of the predicted values to the human ratings

```{r}

# function for correlation data
k_fold_stats <- function(model){           # takes the trained regression model
  per_fold_stats <- model$pred %>%         # correlation stats per fold
    mutate(residuals = obs - pred) %>%
    group_by(Resample) %>% 
    summarise(RMSE = sqrt(mean(residuals^2)),
              R2 = cor(obs, pred)^2,
              MEA = mean(abs(residuals)),
              rho = cor.test(x = obs, y = pred, method = 'spearman')$estimate,
              rho_p = cor.test(x = obs, y = pred, method = 'spearman')$p.value)
  
  # list(per_fold = per_fold_stats, 
  #      summary = per_fold_stats %>% 
  #        summarise(#RMSE_mean = mean(RMSE), RMSE_sd = sd(RMSE),
  #           #R2_mean = mean(R2), R2_sd = sd(R2),
  #           #MEA_mean = mean(MEA), MEA_sd = sd(MEA),
  #           rho_mean = mean(rho), rho_sd = sd(rho),
  #           rho_max = max(rho), rho_min = min(rho))
  #      )
  per_fold_stats %>%
    summarise(rho_mean = mean(rho), rho_sd = sd(rho),
              rho_max = max(rho), rho_min = min(rho))
}
```


**3. Regression Analysis**

Splitting the data into different combination dataframes for the regression analyses.
Models for valence and arousal are kept separate (all models for both val and aro).
Training the models with linear and lasso regression (both tested for all models).
Models: word embeddings - Google and ftext, 
        image embeddings - Resnet and Effnet (both named 'vec' in original dfs, therefore renamed to 'vec_name1', 'vec_name2'),
        multimodal embeddings - all combinations (G+Resnet, G+Effnet, F+Resnet, F+Effnet)
```{r}


models <- list(
  "aro" = list(                    # variables for different arousal models
  "text" = c("AH", "AL"),
  "eff" = c("AH.eff", "AL.eff"),
  "res" = c("AH.res", "AL.res")
  ),
  "val" = list(                    # variables for different valence models
    "text" = c("VH","VL"),
    "eff" = c("VH.eff","VL.eff"),
    "res" = c("VH.res","VL.res")
  )
)

rating_types <- c("val" = "V.Mean.Sum", "aro" = "A.Mean.Sum")     # valence ratings from column 'V.Mean.Sum', arousal from 'A.Mean.Sum'

model_combs = expand.grid(c("glmnet", "lm"), c("val","aro"), c("","google", "ftext"), c("", "eff", "res")) %>%   # automatically create all combinations of elements (unimodal models only take one type embdding, therefore empty added)
  filter(! (Var3 == "" & Var4 == "")) %>%                      # remove case of no word and no image vector 
  mutate_each(as.character) %>%                                # correct type
  rowwise() %>% 
  group_map(~list(model = .$Var1, rating_type = .$Var2, text_vec = .$Var3, image_vec = .$Var4)) # seperates dataframe into list with element for each row, and sublists with element for each column

get_data_by_name <- function(rating_type, vec_type){
  if(vec_type == "") return()
  if(vec_type %in% c("google", "ftext")){           # if text model chosen
    df_names <- models[[rating_type]][["text"]]     # from var 'text'   
    map_dfr(df_names, ~get(.) %>% 
            select(Word, rating_types[rating_type], starts_with(vec_type))  # get word embeddings 
            ) %>%
      return()
  } else {
    df_names <- models[[rating_type]][[vec_type]]
    map_dfr(df_names, ~get(.) %>%
            select(Word, rating_types[rating_type], starts_with("vec")) # otherwise get image embeddings
            ) %>%
      return()
  }
}

# Test:
# walk(model_combs, function(combs){
#   print(combs)
#   map_dbl(c(combs$text_vec, combs$text_vec), 
#           ~get_data_by_name($rating_type, .) %>% nrow() %>% ifelse(is.null(.), 0, .) 
#           ) %>%
#     print()
# })

combine_models <- function(case){
  rating_type <- case$rating_type
  vec_name1 <- case$text_vec                                                  # source of the embedding (e.g. google, res, etc.)
  vec_name2 <- case$image_vec
  col_name1 <- ifelse(vec_name1 %in% c("google", "ftext"), vec_name1, "vec")  # name of the column containing the embedding vector
  col_name2 <- ifelse(vec_name2 %in% c("google", "ftext"), vec_name2, "vec")
  
  if("" %in% c(vec_name1, vec_name2)){                                        # unimodal case
    c(vec_name1, vec_name2)[which(c(vec_name1, vec_name2) != "")] %>% 
      get_data_by_name(rating_type, .) %>%
      (function(df)
        normalise_embedding(df, starts_w = c(col_name1, col_name2)[which(c(vec_name1, vec_name2) != "")]) %>%   # normalise either word or image embedding depending of the model
          left_join(select(df, Word, !!rating_type), by = "Word")
      ) %>%
      return()
  } else {                                                                    # multimodal case
    get_data_by_name(rating_type, vec_name1) %>%
      (function(df)
        normalise_embedding(df, starts_w = col_name1) %>%          # normalise the word embedding
          left_join(select(df, Word, !!rating_type), by = "Word")
      ) %>%
      (function(df)
        get_data_by_name(rating_type, vec_name2) %>%               # normalise the image embedding
          normalise_embedding(., starts_w = col_name2) %>%
          left_join(df, by = "Word")
      ) %>%
      return()
  }
}

# Test:
# map(model_combs, ~combine_models(.) %>% (function(df) c(nrow(df), ncol(df))))


# training the regression models
train_model <- function(case, data){
  reg_type = case$model
  rating_type = case$rating_type
  set.seed(125)
  train.control <- trainControl(method = "cv", number = 10, savePredictions = "all")  #10-fold cross-validation
  if(reg_type == "glmnet"){                    # lasso regression
    train(as.formula(paste(rating_type,"~ .")), 
          data = data, 
          method = "glmnet",
          trControl = train.control,
          tuneGrid = expand.grid(alpha = 1, lambda = seq(0.001, 0.1, by = 0.0002))) %>%
      return()
  } else {                                    # linear regression
    train(as.formula(paste(rating_type,"~ .")), 
          data = data, 
          method = "lm",
          trControl = train.control) %>% 
      return()
  }
}

# Test:
# model_combs[[2]] %>% (function(case){
#   combine_models(case) %>% 
#     select(-Word) %>%
#     train_model(case = case, data = .) %>% 
#     k_fold_stats() %>%                         # get k-fold results
#     mutate(model = case$model, rating = case$rating_type, 
#            text_vec = case$text_vec, image_vec = case$image_vec)
#   })

all_model_stats <- map_dfr(model_combs, function(case){
  combine_models(case) %>%
    select(-Word) %>%
    train_model(case = case, data = .) %>%
    k_fold_stats() %>%
    mutate(model = case$model, rating = case$rating_type,
           text_vec = case$text_vec, image_vec = case$image_vec)
        })

all_model_stats %>% write_csv("./all_model_stats.csv")  # save results

```

**4. Additional test**

Look at words, their original ratings, the predicted ratings and their residuals in the best multimodal model (Google+Resnet, using lasso)
```{r}

good_models <- 
  map(list(
    list(model = "glmnet", rating_type = "val", text_vec = "google", image_vec = "res"),  # for valence
    list(model = "glmnet", rating_type = "aro", text_vec = "google", image_vec = "res")   # for arousal
    ),
     (function(case){
       combine_models(case) %>% 
         select(-Word) %>%
         train_model(case = case, data = .)# %>% 
         # k_fold_stats() %>%
         # mutate(model = case$model, rating = case$rating_type, 
         #       text_vec = case$text_vec, image_vec = case$image_vec)
    })
  )


multimodal_data <- 
  map(list(
    list(model = "glmnet", rating_type = "val", text_vec = "google", image_vec = "res"),
    list(model = "glmnet", rating_type = "aro", text_vec = "google", image_vec = "res")
    ), combine_models
  )

good_models[[1]] %>% class() 

multimodal_data[[2]] %>%  # table from arousal model
  (function(df)
    df %>% select(Word, aro) %>% 
    bind_cols("pred" = predict(good_models[[2]], type = "raw",
                               newdata = df))
  ) %>% 
  mutate(delta = abs(aro - pred)) %>%   # the residuals
  left_join(valar %>% select(Word, aro = A.Mean.Sum, val = V.Mean.Sum)) %>%   # get both ratings
  arrange(desc(delta)) %>%
  head(10) %>% 
  transmute(Word, val, aro, predicted_aro = round(pred,2), residual = round(delta,2)) %>%  # columns from word, both ratings, the predicted arousal rating, residual
  write_delim("worst_aro.txt", delim = " amp ", quote = "none", eol = " \\\\\n")

multimodal_data[[1]] %>%   # table from valence model
  (function(df)
    df %>% select(Word, val) %>% 
    bind_cols("pred" = predict(good_models[[1]], type = "raw",
                               newdata = df))
  ) %>% 
  mutate(delta = abs(val - pred)) %>% 
  left_join(valar %>% select(Word, aro = A.Mean.Sum, val = V.Mean.Sum)) %>%
  arrange(desc(delta)) %>%
  head(10) %>% 
  transmute(Word, val, aro, predicted_val = round(pred,2), residual = round(delta,2)) %>% 
  write_delim("worst_val.txt", delim = " amp ", quote = "none", eol = " \\\\\n")
  


```









